Selective stock enhancement of an exploited species undermines community stability

### 2021-12-06 rmd/manuscript.Rmd
edit: manuscript

---
title: "Selective enhancement of an exploited species undermines community stability"
output:
  word_document:
    reference_docx: word_format.docx
csl: pnas.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: reference.bib
---

@@

Anthropogenic influences have caused global biodiversity crisis, such that human interventions are critical to sustainable management of natural resources. Captive breeding - DEFINE CAPTIVE BREEDING - is a common method to conserve or enhance wild populations of diverse plant and animal taxa. Although there is growing recognition that the selective enhancement with captive-bred organisms entails ecological risks (e.g., the accumulation of deleterious alleles), this method is still pervasive in conservation and fisheries. In fisheries for example, XXX have been spent annually to support hatchery programs globally. This is perhaps because of the "myth" that numerical benefits of released individuals may exceed the associated risks. The current debate, however, overlooks the fact that we have rarely assessed the community-wide impact of selective enhancement. Species are all embedded in the complex web of interacting organisms, and excessive release of particular species may have unforeseen consequences for ecological communities in the long term. Evidence for this hypothesis is lacking, however.

Here, we use a simple, general theory to show that the selective enhancement of a particular species undermines long-term community stability, which we define here as the relative size of fluctuations in total community density over time. The present study further demonstrates the relevance of this general theory to natural systems by illustrating its congruence with stream fish communities, in which an exploited species, masu salmon (*Oncorhynchus masou masou*), is supplemented by hatchery programs. Our findings paint a bleak picture for this form of ecosystem management, reinforcing the recurrent calls for alternative strategies.

We use a multispecies Ricker model to simulate community dynamics with selective enhancement of a constituent species. Specifically, population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$ (intraspecific competition if $i=j$; $\alpha_{ii}=1$), $K_i$ the carrying capacity, $R_t$ the number of released individuals for the enhancement of species (species 1; set constant as $R_t = R$ for simplicity), and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations. The parameter $\phi_i$ takes a value of one when $i=1$ (otherwise zero) to represent the reproductive contribution of released individuals to the next generation of the enhanced species. XX parameter combinations were used to represent ecological scenarios with different strength of competition and XX. Under each simulation scenario, we ran 1400 time steps of 1000 independent communities (i.e., simulation replicates) with varying $R$, of which the last 1000 time steps were used to obtain the coefficient of variation, mean, and standard deviation of community density ($\sum_i^SN_j$). Temporal means and standard deviations were calculated for the enhanced species $N_1$ and the unenhanced species $\sum_{i,i \ne 1}^S N_i$ separately to infer mechanisms that underlie community-level responses.

Our model predicted a destabilizing effect of selective enhancement on ecological communities. Increasing numbers of released individuals $R$ led to greater temporal fluctuations of communities as illustrated by increased CV (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced strong overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals (increased reproducing individuals). Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, the reduced mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. Combined, the long-term average of the total community density decreased more sharply than individual species groups with increased numbers of released individuals (XX). The SDs showed a similar trend, but the relationship was flatter than the mean at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV. These patterns were qualitatively similar across ecological scenarios.

Importantly, the reduced mean of community density may arise as long as humans control the number of releases. The stability of ecological communities rests on internal density regulation that allows for stable coexistence of competing species. For example, significant niche difference in resource use ($\alpha_{ij, i \ne j} < 1$) enables competing species to grow from low population densities because the numerically dominant species undergoes stronger intraspecific competition. However, the number of release is not subject to density-dependent feedback. The excessive individuals released by humans impose competition that would otherwise not exist and undermine the self-regulation process of stable communities. This argument is supported by the fact that the destabilizing effect of selective enhancement manifested even with weak interspecific competition that allows for stable coexistence of competing species. Perhaps, the only exception would be a scenario in which the species of interest is on the verge of local extinction with little or no density regulation (i.e., very large carrying capacity).

EMPIRICAL ANALYSIS

@@

We developed a multispecies Ricker model to simulate community dynamics with selective enhancement of a particular species. Population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient between species $i$ and $j$ (intraspecific competition if $i=j$; $\alpha_{ii}=1$), $K_i$ the carrying capacity, $R_t$ the number of released individuals for the enhancement of species 1, and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations. The parameter $\phi_i$ takes a value of one when $i=1$ (otherwise zero) to represent the reproductive contribution of released individuals to the next generation of the enhanced species. The species-specific intrinsic growth rates and competition coefficients were drawn randomly from a uniform and exponential distribution, respectively, to allow for interspecific variation as $r_i \sim Unif(0.5, r_{max})$ and $\alpha_{ij, i \ne j} \sim Exp(\mu_{\alpha})$. The number of released individuals was assumed to be constant across time steps ($R_t = R$). Environmental stochasticity was drawn from a normal distribution as $\epsilon_{i,t} \sim Normal(\mu_{\epsilon}, \sigma_{\epsilon}^2)$

@@

Under each parameter combination, we ran 1400 time steps of 1000 independent communities (i.e., simulation replicates) with the gradient of the number of released individuals $R$ (0 - 500). We initialized the community with populations of each species drawn from a poisson distribution with a mean of five. We repeated the seeding procedure every 10 time steps over the first 200 time steps to allow species to establish populations and reach an equilibrium without (initialization period). After the initialization period, we released $R$ individuals of the enhanced species every time step over the next 200 time steps to reach a new equilibrium with selective enhancement (burn-in period). We saved the last 1000 time steps to obtain the relative temporal fluctuations (i.e., the coefficient of variation) of the whole community ($\sum_i^SN_j$), the enhanced species ($N_1$), and the unenhanced species ($\sum_{i,i \ne 1}^S N_i$). We also calculated temporal means and standard deviations for each species group. Values of simulation parameters were summarized in XXX. We performed simulations using R version XXX.

>>>>

---
title: "Selective stock enhancement of an exploited species undermines community stability"
output:
  word_document:
    reference_docx: word_format.docx
csl: pnas.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: reference.bib
---

@@

Anthropogenic influences have caused global biodiversity crisis, such that human interventions are critical to sustainable management of natural resources. Captive breeding is a common practice to conserve or enhance wild populations of diverse plant and animal taxa. Although there is growing recognition that stock enhancement with captive-bred organisms entails ecological risks (e.g., the accumulation of deleterious alleles), this method is still pervasive in conservation and fisheries. In fisheries for example, XXX have been spent annually to release billions of hatchery-reared individuals globally. This is perhaps because of the "myth" that numerical benefits of released individuals may exceed the associated risks. The current debate, however, overlooks the fact that we have rarely assessed the community-wide impact of selective stock enhancement. Species are all embedded in the complex web of interacting organisms, and excessive release of particular species may have unforeseen consequences for ecological communities in the long term. Evidence for this hypothesis is lacking, however.

Here, we use a simple, general theory to show that the selective stock enhancement undermines long-term community stability, which we define here as the relative size of fluctuations in total community density over time. The present study further demonstrates the relevance of this general theory to natural systems by illustrating its congruence with Japanese stream fish communities, where XXX of hatchery masu salmon (*Oncorhynchus masou masou*) are released for fisheries, recreation, and conservation purposes. Our findings paint a bleak picture for this form of ecosystem management, reinforcing the recurrent calls for alternative strategies.

We use a multispecies Ricker model to simulate community dynamics with selective stock enhancement of a constituent species. Specifically, population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$ (intraspecific competition if $i=j$; $\alpha_{ii}=1$), $K_i$ the carrying capacity, $R_t$ the number of released individuals for the enhancement of species 1 (set constant as $R_t = R$ for simplicity), and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations. The parameter $\phi_i$ takes a value of one when $i=1$ (otherwise zero) to represent the reproductive contribution of released individuals to the next generation of the enhanced species. XX parameter combinations were used to represent ecological scenarios with different strength of competition and XX. Under each simulation scenario, we ran 1400 time steps of 1000 independent communities (i.e., simulation replicates) with varying $R$, of which the last 1000 time steps were used to obtain the coefficient of variation, mean, and standard deviation of community density ($\sum_i^SN_j$). Temporal means and standard deviations were calculated for the enhanced species $N_1$ and the unenhanced species $\sum_{i,i \ne 1}^S N_i$ separately to infer mechanisms that underlie community-level responses.

Our model predicted a destabilizing effect of selective stock enhancement on ecological communities. Increasing numbers of released individuals $R$ led to greater temporal fluctuations of communities as illustrated by increased CV (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced strong overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals (increased reproducing individuals). Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, the reduced mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. Combined, the long-term average of the total community density decreased more sharply than individual species groups with increased numbers of released individuals (XX). The SDs showed a similar trend, but the relationship was flatter than the mean at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV. These patterns were qualitatively similar across ecological scenarios.

Importantly, the reduced mean of community density may arise as long as humans control the number of releases. The stability of ecological communities rests on internal density regulation that allows for stable coexistence of competing species. For example, significant niche difference in resource use ($\alpha_{ij, i \ne j} < 1$) enables competing species to grow from low population densities because the numerically dominant species undergoes stronger intraspecific competition. However, the number of release is not subject to density-dependent feedback. The excessive individuals released by humans impose competition that would otherwise not exist and undermine the self-regulation process of stable communities. This argument is supported by the fact that the destabilizing effect of selective stock enhancement manifested even with weak interspecific competition that allows for stable coexistence of competing species. Perhaps, the only exception would be a scenario in which the species of interest is on the verge of local extinction with little or no density regulation (i.e., very large carrying capacity).

## Empirical evidence

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of selective stock enhancement of masu salmon on long-term stability of stream fish communities. Masu salmon represents one of the most valuable freshwater resources in Hokkaido Island, Japan, and XX leads extensive hatchery programs across the island. In protected watersheds where any form of exploitation is prohibited (including recreational angling), long-term monitoring has been conducted with consistent sampling methods over the past two decades along with official records of the number of fish released. Therefore, the data set the stage for a "natural experiment" to test the theoretical predictions. In this study, we assembled time-series data of stream fish communities at XX sites within XX protected watersheds (see Methods for selection criterion). Using hierarchical Bayesian models, we quantified the relationship between community dynamics and the number of fish released while accounting for other potential confounding factors.

As our theory predicts, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with greater numbers of fish released. The positive relationship between the CV and the number of fish released was strongly supported by a hierarchical linear model, in which the probability of the regression coefficient being positive was XXX. This was associated with the reduced long-term averages of population densities of enhanced (masu salmon) and unenhanced fish groups. Both fish groups showed low population densities with high numbers of fish released (STATS), although statistical support for this relationship in the individual fish groups was weak to moderate (STATS). In the meantime, the SDs of population densities had vague relationships with the number of fish released, and neither positive or negative relationship was supported by statistical analysis.

Our study system satisfies key theoretical assumptions that are required for the destabilizing effect of selective stock enhancement to emerge. The first assumption is that released fish compete with wild fish. This type of competition is very likely to occur in the study streams. Masu salmon typically spend one year in freshwater before migrating to the ocean, and a wealth of evidence suggests that this species compete with other species of stream fish communities. In particular, masu salmon raised in hatchery environments are more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild. Second, our theory predicts the emergence of destabilizing effect when the number of fish released is externally controlled. This is exactly what was observed in our study system. In the protected watersheds, there was no apparent congruence between the number of fish released and the population density of masu salmon (XX). Instead, these released fish are likely controlled by socioeconomic factors, such as budget allocation to hatchery programs in this region. As such, released fish are probably "excessive" given the density-dependent regulation of natural communities, imposing intense resource competition that otherwise would not exist.

Genetic effect XXX

## Implications

Although captive breeding programs carefully consider the fitness of captive-bred individuals in the wild, the debate often ignores the capacity of recipient ecosystems to support biodiversity. The results of this study suggest that the ignorance of the ecosystem capacity is critical, undermining the long-term stability of natural communities. Our research has two important implications to improve captive breeding programs. First, release of captive-bred individuals should be avoided when strong density-dependence operates in the recipient ecosystem. Second, we should prioritize the preservation and restoration of recipient ecosystems, which are rapidly degraded due to myriads of human influences. Without this integrative management framework that recognizes the fundamental process of biodiversity maintenance, selective stock enhancement will never be effective but will impair biodiversity.

@@

We developed a multispecies Ricker model to simulate community dynamics with selective stock enhancement of a particular species. Population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient between species $i$ and $j$ (intraspecific competition if $i=j$; $\alpha_{ii}=1$), $K_i$ the carrying capacity, $R_t$ the number of released individuals for the stock enhancement of species 1, and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations. The parameter $\phi_i$ takes a value of one when $i=1$ (otherwise zero) to represent the reproductive contribution of released individuals to the next generation of the enhanced species. The species-specific intrinsic growth rates and competition coefficients were drawn randomly from a uniform and exponential distribution, respectively, to allow for interspecific variation as $r_i \sim Unif(0.5, r_{max})$ and $\alpha_{ij, i \ne j} \sim Exp(\mu_{\alpha})$. The number of released individuals was assumed to be constant across time steps ($R_t = R$). Environmental stochasticity was drawn from a normal distribution as $\epsilon_{i,t} \sim Normal(\mu_{\epsilon}, \sigma_{\epsilon}^2)$

@@

Under each parameter combination, we ran 1400 time steps of 1000 independent communities (i.e., simulation replicates) with the gradient of the number of released individuals $R$ (0 - 500). We initialized the community with populations of each species drawn from a poisson distribution with a mean of five. We repeated the seeding procedure every 10 time steps over the first 200 time steps to allow species to establish populations and reach an equilibrium without (initialization period). After the initialization period, we released $R$ individuals of the enhanced species every time step over the next 200 time steps to reach a new equilibrium with selective stock enhancement (burn-in period). We saved the last 1000 time steps to obtain the relative temporal fluctuations (i.e., the coefficient of variation) of the whole community ($\sum_i^SN_j$), the enhanced species ($N_1$), and the unenhanced species ($\sum_{i,i \ne 1}^S N_i$). We also calculated temporal means and standard deviations for each species group. Values of simulation parameters were summarized in XXX. We performed simulations using R version XXX.
